Overview

### 2024-01-16 analysis/001_overview.Rmd
Start describing subject

Next, I focus on the subject distribution.

@@

Here's a table. I applied fractional counting because of large overlaps.

@@

```{r subject_plots}
subject_trend_plot <- top_level_by_year_oa |>
  filter(top_level != "Multidisciplinary") |>
  filter(!cr_year %in%  c("2017", "2023")) |>
  select(cr_year, top_level, contains("frac"),-oa_frac) |>
  pivot_longer(
    cols = c(ta_oa_frac, no_ta_fractions),
    names_to = "type",
    values_to = "value"
  ) |>
  mutate(prop = value / articles_frac) |>
  ggplot(aes(
    gsub("^20", "'", as.character(cr_year)),
    prop,
    fill = type,
    group = type
  )) +
    geom_area(color = "#000000",
            linewidth = 0.15) +
  facet_wrap( ~ top_level, nrow = 1) +
  scale_fill_manual(
    "Hybrid OA",
    values = c(
      no_ta_fractions = "#b3b3b3a0",
      ta_oa_frac = "#56B4E9"
    ),
    labels = c(no_ta_fractions = "Other", ta_oa_frac = "Agreement"),
    guide = guide_legend(reverse = TRUE)
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05)),
    labels = scales::percent_format(),
    breaks = c(0.05, 0.1, 0.15, 0.2, 0.25)
  ) +
  theme_minimal() +
  labs(y = "Hybrid OA Uptake", x = NULL) +
  theme(
    legend.position = "top",
    legend.justification = "left",
    axis.text.x = element_blank()
  )

@@

@@

And here's my country analysis. First top 20

@@

>>>>

@@

Table presents a high-level overview of hybrid open access by AJCS subject area using fractionalised counting to account for journals belonging to more than one category. 
Between 2018 and 2022, most hybrid journals with at least one open articles could be attributed to the social sciences including the humanities. 
However, these journals published the fewest number of articles, whereas physical sciences journals recorded most articles (n = , 44%), followed by the health sciences and the life sciences. 
In terms of open access,  physical sciences journals accounted for more than one third of articles published in the five-years period, followed by the health science, the social sciences and the life sciences.

@@

```{r subject_plots}
subject_trend_plot <- top_level_by_year_oa |>
  filter(top_level != "Multidisciplinary") |>
  filter(!cr_year %in%  c("2017", "2023")) |>
  select(cr_year, top_level, contains("frac"),-oa_frac) |>
  pivot_longer(
    cols = c(ta_oa_frac, no_ta_fractions),
    names_to = "type",
    values_to = "value"
  ) |>
  mutate(prop = value / articles_frac) |>
  ggplot(aes(
    gsub("^20", "'", as.character(cr_year)),
    prop,
    fill = type,
    group = type
  )) +
    geom_area(color = "#000000",
            linewidth = 0.15) +
  facet_wrap( ~ top_level, nrow = 1) +
  scale_fill_manual(
    NULL,
    values = c(
      no_ta_fractions = "#b3b3b3a0",
      ta_oa_frac = "#56B4E9"
    ),
    labels = c(no_ta_fractions = "Other", ta_oa_frac = "Agreement"),
    guide = guide_legend(reverse = TRUE)
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05)),
    labels = scales::percent_format(),
    breaks = c(0.05, 0.1, 0.15, 0.2, 0.25)
  ) +
  theme_minimal() +
  labs(y = "Hybrid OA Uptake", x = NULL) +
  theme(
    legend.position = "top",
    legend.justification = "left",
    axis.text.x = element_blank()
  )

@@

Figure presents the relative growth of hybrid open access by subject area between 2018-2022. In particular, journals in the social sciences accounted for the strongest growth by from % to %, followed by the Life Science , Health Science and Physical Sciences. 
This growth in the social sciences can be largely attributed to transformative agreements, which are prevalent across journals in this field. In 2022, 75% , (median).
However, it is noteworthy that hybrid open access through transformative agreements plays a comparable lesser role in the life sciences and health sciences. 
In these two subject areas, only about half of the open access articles could be linked to transformative agreements, both overall and on median average across journals.
In contrast to the Life Science , Health Science, the majority of physical science journals, recorded an above-average open access uptake through transformative agreements.

@@

@@

### Supplements

```{r country_by_publisher}
country_jn_all <- hoaddata::jn_aff |>
  filter(!cr_year %in% c(2017, 2023), issn_l != "0027-8424") |>
  # only with cc
  filter(issn_l %in% hoaddata::cc_articles$issn_l) |>
  distinct(issn_l, cr_year, country_code, articles_total) |>
  group_by(cr_year, country_code, issn_l) |>
  summarise(all = sum(articles_total))

country_jn_oa <- hoaddata::cc_articles |>
  filter(!cr_year %in% c(2017, 2023), issn_l != "0027-8424") |>
  group_by(cr_year, country_code, issn_l) |>
  summarise(oa = n_distinct(doi))

country_codes <- readr::read_csv(here::here("data", "ror_country_codes.csv"))
country_jn_ta <-  ta_oa_inst |>
  filter(ta_active == TRUE, !is.na(cc), issn_l != "0027-8424", !cr_year %in%  c("2017", "2023")) |>
  left_join(country_codes) |>
  group_by(cr_year, country_code, issn_l) |>
  summarise(ta_oa = n_distinct(doi))

country_jn_df <- country_jn_all |>
  left_join(country_jn_oa, by = c("cr_year", "country_code", "issn_l")) |>
  left_join(country_jn_ta, by = c("cr_year", "country_code", "issn_l")) |>
  distinct() |>
  ungroup() |>
  mutate(cr_year = as.character(cr_year)) |>
  mutate(across(c(oa, ta_oa), .fns = ~replace_na(.,0))) |>
  filter(!is.na(country_code)) |>
  mutate(
    country_name = countrycode::countrycode(country_code, origin = "iso2c", destination = "country.name")
  )

## add publishers
country_by_publisher_df <- hoaddata::jct_hybrid_jns |>
  distinct(issn_l, esac_publisher) |>
  inner_join(country_jn_df, by = "issn_l") |>
  # Publisher encoding
   mutate(
    publisher = case_when(
      esac_publisher == "Elsevier" ~ "Elsevier",
      esac_publisher == "Springer Nature" ~ "Springer Nature",
      esac_publisher == "Wiley" ~ "Wiley",
      is.character(esac_publisher) ~ "Other"
    )
  ) |>
  mutate(publisher =
           forcats::fct_relevel(
             publisher,
             c("Elsevier", "Springer Nature", "Wiley", "Other") 
           )) |>
  # Country encoding
   mutate(cat = case_when(
    country_code %in% oecd$country_code ~ "OECD",
    country_code %in% c("BR", "CN", "IN", "RU", "ZA") ~ "BRICS",
    .default = "Other"
  )) |>
  # Summary
  group_by(cr_year, publisher, cat) |>
  summarise(all = sum(all),
            oa = sum(oa),
            ta_oa = sum(ta_oa), .groups = "drop") |>
  mutate(closed = all - oa,
         oa_other = oa - ta_oa)

# Plot
country_by_publisher_df |>
  pivot_longer(cols = c(ta_oa, oa_other, closed), names_to = "oa_type") |>
  ggplot(aes(gsub("^20", "'", as.character(cr_year)), value, group = oa_type, fill = oa_type)) +
  geom_area(color = "#000000",
            linewidth = 0.15) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), labels =  scales::number_format(big.mark = ",")) +
  facet_grid(cat ~publisher) +
  scale_fill_manual("Hybrid OA", values = c(oa_other = "#b3b3b3a0", ta_oa = "#56B4E9", closed = "#e1e1e1a0"),
                    labels = c(oa_other = "Other", ta_oa = "Agreement", closed = "NA"),  guide = guide_legend(reverse = TRUE)) +
  theme_minimal() +
  labs(x = "Publication year", y = "Total articles")
```
